---
title: "Prison System Vs Race" 
author: "Truc Quynh Nguyen"
date: 11-22-2022
output: html_document
---

```{r echo=FALSE, eval=TRUE, message=FALSE}
source("../source/a4-helpers.R")
source("../source/analysis.R")
```

## Abstract 
This report will showcase the prision system in the US in the past 50 years. We will take a look at how different genders are in the systems and also which states had the most imprisonment

```{r echo=FALSE, eval=TRUE, message=FALSE}

```



```{r echo=TRUE, eval=TRUE, message=FALSE}

```

## Introduction 
In this project, I will showcase how different races has been affected by the jail system. We will take a look at data and see if there is a certain race group that is being treated more unfairly or not verus other races. To beginning, we know that jail is a place for people has commited crimes. But are all the cases fair? or are certain groups being targeted just for the color of their skin? The direct stakeholders in this project will refer to those people of color who have been affected. Indirected stakeholders will refer to those in their communities that have been affected by what is happening. 

## Data Summary 
From the dataset, I found out that California has the most arrest. I though that this was an intresitng fact because it tells us about the system in California. Another fact I found out is that the highest amount of black people ever recored to be arrested within a year was 13143 people. Looking at this, I was suprised because that is alot of people within just one race group. The last thing I wanted to find out from the data set was how many white people were arrested within one year. The number turned out to be 7036 people within one year. It is ovibous that there is a large differnce between the two races. This might show there is inequality within the system. 


<br />

## Growth of the U.S. Prison Population

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplotly(Plot)
```
I did a bar chart to visualize how the jail population has increased since 1970 till 2020. This is a good way to graph because it shows how the numbers have increased drastically. Looking at the bar chart, we can see that the jail poluation began to rise starting in 1980. After 1980, the numbers began to rise. Looking at the bar chart, we can also see that the highest number was in 2008. 

## Growth of Prison Population by State 
```{r echo=FALSE, message=FALSE, warning=FALSE}
states <- c("WA", "OR", "CA")
ggplotly(plot_jail_pop_by_states(states))
```
This is a line chart that shows the number of jail populations in the states of Calionia, Oregon, and Washington. We can see that Califorina has the highest number of inmates. Looking at the line chart, there is an obvious difference between the states. We can see that Orgeon and Washington has lowest numbers then Califorina. 

## Patterns of inequality between male and female. 
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplotly(sex_plot())
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplotly(female_sex_plot())
```
These are bar charts that compare the female and male populations between 1970 and 2020. As can be seen, males have always had a higher number of jail population. We can see that the highest number of male inmates were in 2008 when females were in 2018. We can also see that female inmate were higher than males near 2017.

